---
output: html_document
---

###**Victor D. Saldaña C.**
######PhD(c) in Geoinformatics Engineering
######Technical University of Madrid (Spain)
######LinkedIn: https://es.linkedin.com/in/victorsalda
######Github: https://github.com/victorsalda
######Rpubs: http://rpubs.com/victorsalda
######email: victorsalda@hotmail.com
######*November, 2016*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Summary.

This is the report of the design of a web page using R Markdown that features a plot created with Plotly package. The "airquality" data set is going to be use

### 2. Load packages.

Let's load the packages needed. If you haven't download them do it first.

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library("plotly","ggplot2","airquality")
airquality<-airquality
```

### 3. Plot created with Plotly.

Let's create the plot a Wind vs Temperature using some basic plotly commands. 

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
plot_ly(airquality,x=wind,y=~Temp,mode="markers",color =~Day, 
        name = "Wind vs Temperture",size=~Ozone)
#the use of tilde "~" is necesary in most veraibles but not clear. More info: 
#https://github.com/ropensci/plotly/issues/765
```
