---
title: "Regression outputs on a grid of ggplot2 plots (facets)"
output: html_document
---
####**Victor D. Saldaña C.**
######PhD(c) in Geoinformatics Engineering
######Technical University of Madrid (Spain)
######https://es.linkedin.com/in/victorsalda
######https://github.com/victorsalda
######*June 2016*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###**Overview.**
####This script allows to add to a group of ggplot2 plots laid out in panels with facets_grid 
####the values of the slope, intercept, R^2 and adjusted R^2 of every plot.

###**Context.**
####Let's suppose you have a data frame with some interesting data and you want 
####to do some regression to fit a linear model with a straight line. Likewise, you want
####to do the regression analysis by sub setting the data. In this case, ggplot2 package
####offers a good set of tools to do it by dividing the date using "facets" that divides 
####the data frame according to the values of certain column. If the column you use to subset
####the data frame have four values you will get four plots with its own regression line. Now,
####if you want to add the regression equation and/or others values to compare them is 
####not possible with gggplot2, at least, in a direct way. Here is an option. 

###**1. Load packages.**
```{r }
#1. Let's load the packages we need. If you haven't download them do it first.
library("ggplot2")
library("datasets")
library("plyr")
```

###**2. Create a function.**
```{r}
#2. We need to create a function. Let's call it "regression". It has only one argument, 
#the data frame (df). Likewise, let's use the "Toothgrowth" dataset from the package 
#“datasets”. This dataset contains 60 observations of "the Effect of Vitamin C on 
#Tooth Growth in Guinea Pigs". There are three variables: 
#1) Len (Tooth Length, numeric).
#2) Supp (Supplement Type, OJ = orange juice or VC = vitamin C, factor).
#3) Dose (Dose levels of vitamin, 0.5, 1, and 2 mg/day, numeric).

#2.1. Load the data and set a data frame (df).
data("ToothGrowth") #Load the dataset "ToothGrowth"
df<-ToothGrowth # Assigning it to a variables. 

#2.2. Create function.
regression=function(df){
        #setting the regression function. 
        reg_fun<-lm(formula=df$len~df$dose) #regression function
        #getting the slope, intercept, R square and adjusted R squared of 
        #the regression function (with 3 decimals).
        slope<-round(coef(reg_fun)[2],3)  
        intercept<-round(coef(reg_fun)[1],3) 
        R2<-round(as.numeric(summary(reg_fun)[8]),3)
        R2.Adj<-round(as.numeric(summary(reg_fun)[9]),3)
        c(slope,intercept,R2,R2.Adj)
        }
```

###**3.Subset the data frame.**
```{r}
#3. Now let's subset the data frame base on the values of the supplement type (supp) and
#let's apply the function "regression" to every data frame. To achieve this
#let's use the "ddply" function from the "plyr". The results in this case is another
#data frame (regressions_data) whose rows are the valeus of the slope, intercept, 
#R_squared and R_squared_adj for every data frame.
regressions_data<-ddply(ToothGrowth,"supp",regression)
colnames(regressions_data)<-c ("supp","slope","intercept","R2","R2.Adj")
```

###**4. Plotting.**[^1][^2]
```{r}
#4. Let's plot all the plots and the values of the regression models. 
qplot(dose, len, data = ToothGrowth, size=I(2))+geom_smooth(method="lm")+
        facet_grid(supp ~ .)+ggtitle("Regressions")+
        geom_label(data=regressions_data, inherit.aes=FALSE, aes(x = 1.2, y = 32,
label=paste("slope=",slope,","," ","intercept=",intercept,","," ","R^2=",R2,","," ","R^2.Adj=",R2.Adj)))
```

[^1]:1. If you don't want a rectangle underneath the text use "geom_text" insted of "geom_label".
[^2]:2. is possible much more customization. Check out the help file of facets_grid, geom_label, etc.